Consistent Polyhedral Surrogates via Embedding

### 2019-05-15 neurips-19/poly-embed.tex
Separated out proofs into its own tex file

\begin{abstract}
  We formalize and study the natural approach of designing convex surrogate loss functions via embeddings, for problems such as classification or ranking which involve a finite number of possible predictions.
  Specifically, we consider surrogate losses given by embedding each prediction as a point in $\reals^d$, and convexifying the loss.
  We prove that this approach is equivalent, in a strong sense, to working with polyhedral (piecewise linear convex) losses.
  We then investigate the dimensionality of the surrogate hypothesis space required for such an embedding, giving a complete characterization of the 1-dimensional case.
  We illustrate our results with modest lower bounds for abstain loss, and conclude with future directions.
\end{abstract}

@@

Let $\Y$ be a finite outcome (label) space, and throughout let $n=|\Y|$.
The set of probability distributions on $\Y$ is denoted $\simplex\subseteq\reals^{\Y}$, represented as vectors of probabilities.
We write $p_y$ for the probability of outcome $y \in \Y$ drawn from $p \in \simplex$.
\jessiet{Do we want to go into the detail of thinking of $p_y$ as $p(Y|X)$ in finite data}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proving (in)consistency via embeddings}
\raf{Notes about the story, partially to myself, partially to Jessie: I think we want to emphasize the proofs of consistency and inconsistency (as the section title reflects), but we also want to illustrate our constructive approach to finding a link.  Basically, most of these examples (except abstain) have the link built in (usually as the sign function, though slightly different for top-k), and they just work around it.  For abstain, we can give alternate link functions, and hopefully show that the L1-based link has a better calibration function.  For the others, it may suggest that it's worth thinking ``outside the box'' (HAH, genious...) for embeddings which are not the hypercube.}

@@

The top $k$ classification problem takes as input a vector $u \in \reals^n$, and observes outcome $y \in [n]$.
The input is sorted in descending order, and the function $t_k(u)$ yields a set of outcomes associated to the first $k$ elements of this sorted input.
For example, if $u = (4,1,2)$, then $t_k(u) = \{0,2\}$ for the first and third outcomes.

We say the top-$k$ loss is
\begin{align}
\ell_{top-k}(u,y) = \begin{cases}
0 & y \in t_k(u)\\
1 & \text{otherwise}
\end{cases}
\end{align}

Yang and Koyejo~\cite{yang2018consistency} create a new surrogate loss that is consistent for the top-$k$ problem, although it should be noted that their surrogate is not convex.
In particular, Yang and Koyejo evaluate two losses given by Lapin et al.~\cite{lapin2016loss, lapin2018analsis}, one of which is shown to be convex, but its consistency is left as an open question.

After our discussion of embeddings, we now have a way of proving consistency: a surrogate $L$ is consistent with respect to a discrete loss $\ell$ \jessie{we still haven't defined consistency... here I'm going by Yang's consistency.  I suppose we still need a proof of this.} if and only if $L$ embeds $\ell$.
\raf{Consistency: it's is about the expected loss over $(x,y)$ pairs, and showing that getting the best surrogate loss should mean getting the best original loss.  Take another look at Theorem 2.4 of the top-k paper.  Losses can certainly be consistent without embedding $\ell$ (even if polyhedral: you can link multiple embedded points to a single discrete report, e.g.\ a convex surrogate for ranking loss can link to the mode), so the ``iff'' is not right.  I think it's worth spending some time thinking about what exactly we need to show to prove inconsistency.  In particular, see if this reasoning checks out:  Suppose we find a distribution $p$ for which an optimal surrogate report is $r$ but $\psi(r)$ is not optimal for $\ell$.  Since we could have a conditional distribution of $p$ for every $x$, so the optimal hypothesis would just be the constant function $h(x) = r$, which achieves the optimal $L$-loss, and when linked, $\psi \circ h(x) = \psi(r)$, which does not achieve the optimal $\ell$-loss.}
We can then prove that the surrogates given by Lapin et al.~\cite{lapin2018analysis} are not consistent in a more suffice manner; we do so by showing the discrete loss that their surrogate embeds is not equal to  $\ell_{top-k}$.

>>>>

\begin{abstract}
We formalize and study the natural approach of designing convex surrogate loss functions via embeddings for problems such as classification or ranking. 
In this approach, one embeds each of the finitely many predictions (e.g. classes) as a point in $\reals^d$, assigns the original loss values to these points, and convexifies the loss in between to obtain a surrogate.  
We prove that this approach is equivalent, in a strong sense, to working with polyhedral (piecewise linear convex) losses.  
Moreover, given any polyhedral loss $L$, we give a construction of a link function through which $L$ is a consistent surrogate for the loss it embeds.  
We illustrate the power of this embedding framework with succinct proofs of consistency or inconsistency of various polyhedral surrogates in the literature.
\end{abstract}

@@

Let $\Y$ be a finite outcome (label) space, and throughout let $n=|\Y|$.
The set of probability distributions on $\Y$ is denoted $\simplex\subseteq\reals^{\Y}$, represented as vectors of probabilities.
We write $p_y$ for the probability of outcome $y \in \Y$ drawn from $p \in \simplex$.
\jessie{Given labeled data of the form $(x,y)$, we abstract away the data input and consider the conditional probabilities $p(Y|X)$, and this is what we denote as $p_y$.}
\jessiet{Do we want to go into the detail of thinking of $p_y$ as $p(Y|X)$ in finite data}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proving (in)consistency via embeddings}\label{sec:examples}
\raf{Notes about the story, partially to myself, partially to Jessie: I think we want to emphasize the proofs of consistency and inconsistency (as the section title reflects), but we also want to illustrate our constructive approach to finding a link.  Basically, most of these examples (except abstain) have the link built in (usually as the sign function, though slightly different for top-k), and they just work around it.  For abstain, we can give alternate link functions, and hopefully show that the L1-based link has a better calibration function.  For the others, it may suggest that it's worth thinking ``outside the box'' (HAH, genious...) for embeddings which are not the hypercube.}

@@

The top $k$ classification problem takes as input a vector $r \in \R = \{r \in \{0,1\}^n \text{ such that } |r| = k\}$, and observes outcome $y \in [n]$

We say the top-$k$ loss $\ell_{top-k}:\R \to \reals$ is
\begin{align}
\ell_{top-k}(u,y) = \begin{cases}
0 & u_y = 1\\
1 & \text{otherwise}
\end{cases}
\end{align}

Yang and Koyejo~\cite{yang2018consistency} create a new surrogate loss that is consistent for the top-$k$ problem, although it should be noted that their surrogate is not convex.
Additionally, Yang and Koyejo evaluate two losses given by Lapin et al.~\cite{lapin2016loss, lapin2018analysis}, one of which is shown to be convex, but its consistency is left as an open question.

After our discussion of embeddings, we now have a way of proving inconsistency: if a surrogate $L$ embeds a discrete loss $\ell$, then $L$ is consistent with respect $\ell$. \jessie{I suppose we still need a proof of this.} 
Moreover, if for some other discrete loss $\ell' \neq \ell$, 
\raf{Consistency: it's is about the expected loss over $(x,y)$ pairs, and showing that getting the best surrogate loss should mean getting the best original loss.  Take another look at Theorem 2.4 of the top-k paper.  Losses can certainly be consistent without embedding $\ell$ (even if polyhedral: you can link multiple embedded points to a single discrete report, e.g.\ a convex surrogate for ranking loss can link to the mode), so the ``iff'' is not right.  I think it's worth spending some time thinking about what exactly we need to show to prove inconsistency.  In particular, see if this reasoning checks out:  Suppose we find a distribution $p$ for which an optimal surrogate report is $r$ but $\psi(r)$ is not optimal for $\ell$.  Since we could have a conditional distribution of $p$ for every $x$, so the optimal hypothesis would just be the constant function $h(x) = r$, which achieves the optimal $L$-loss, and when linked, $\psi \circ h(x) = \psi(r)$, which does not achieve the optimal $\ell$-loss.}
\jessie{I think what we want to say is the following: if $L$ embeds $\ell$, then $L$ is consistent with respect to $\ell$.  Moreover, if $\ell \neq \ell'$ and $L$ embeds $\ell$, then $L$ does not not embed $\ell'$.  Can we take this and say something about $L$ not being consistent wrt $\ell'$?}
We can then prove that the surrogates given by Lapin et al.~\cite{lapin2018analysis} are not consistent in a more suffice manner; we do so by showing the discrete loss that their surrogate embeds is not equal to  $\ell_{top-k}$.
